### 2019-02-20 paper-kaggle.Rmd
LR added

---
title: Kaggle-in-class Data Challenges Can Boost Student Learning

# to produce blinded version set to 1
blinded: 1

authors: 
  - name: Julia Polak
    affiliation: Department of Statistics, University of Melbourne
  - name: Dianne Cook
    affiliation: Department of Econometrics and Business Statistics, Monash University

header-includes:
    - \usepackage{setspace}\doublespacing

keywords:
- instructional technology
- statistical modeling
- data science
- statistics education
- data mining

abstract: |
  Kaggle is a data modeling competition service, where participants compete to build a model with lower predictive error than other participants. Several years ago they released a reduced service that enables instructors to run competitions in a classroom setting. This paper describes the results of an experiment to determine if participating in a predictive modeling competition enhances learning. The evidence suggests it does. In addition, students were surveyed to examine if the competition improved engagement and interest in the class.

bibliography: bibliography.bib
output: rticles::asa_article
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo=FALSE, 
  message=FALSE, 
  warning=FALSE, 
  kfigr.link=TRUE, 
  kfigr.prefix=TRUE, 
  cache=FALSE, 
  fig.cap=TRUE, 
  fig.height=3)
```

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(knitr)
library(bookdown)
library(knitcitations)
library(RefManageR)
library(tidyverse)
#library(ggpubr)
library(readxl)
library(gridExtra)
library(forcats)

options("citation_format" = "pandoc")

BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic')
```

# Introduction

Kaggle [@kaggle] is well-known for the data competitions, some richly funded. It provides a platform for predictive modelling and analytics competitions where participants compete to produce the best predictive model for a given data set. One of the well-known competitions is the ``Titanic: Machine Learning from Disaster'' in which, mainly the beginners, can practice their skills to predict survival on the Titanic and get familiar with the Kaggle platform and some of the machine learning techniques.

*Di any``famous'' data competition examples ???*

The competitors participating in these competitions have different motivations, but it is mainly fame and financial reward for the winner. 
Occasionally, inventors of new methods and algorithms use Kaggle competitions as a platform to prove the superiority of their methods. A famous example are Geoffrey Hinton and George Dahl with deep neural networks [@HintonDahi12] and Tianqi Chen and Carlos Guestrin with XGBoost [@ChenXGBoost].

Individuals and organizations that contribute the data and open it for a competition gain expertise from many professionals at once. This increases the chance for an improved and innovative solution to their problem.

In 2015, Kaggle InClass was introduced, as a self-service platform to conduct competitions. These competitions can be private, limited to members of a university course, and are easy to setup. This paper examines the educational benefits of conducting predictive modeling competitions in class on performance, engagement and interest.

<!--
Background literature:
- Data competition
- Competition and pedagogy
- Gold standard (randomized control studies) for evaluating new pedagogy
- active learning
-->

In the past few years, educational community started to collect positive evidence on including a competition in the classroom (see for example @Calnon12, @Canada15 and @VanNuland15).
The most comparable to our experiment is the educational experiment conducted by @VanNuland15. 
The authors found significant improvement in student's exam marks due to taking part in the competition. The experiment (crossover design) was conducted in a similar class size ($n=67$). However, these were undergraduate students that took the human anatomy subject at Western University.

The competition is an example of active learning. A large body of literature has been published on the latter topic. @Prince04
surveyed the literature and found that all forms of active learning have positive effect on the learning experience and student's achievements. The magnitude of the effect of different approaches, though, varies. What's more, @Freeman14 examined 158 studies published in about 50 STEM educational journals. The authors found that student's exam marks increased by almost half SD under active learning. Moreover, students in classes with traditional lecturing were 1.5 times more likely to fail than their peers in classes with active learning.

A competition, like any other assessment method has its advantages and disadvantages. It brings the 'game' feeling, increases the interest level among students and motivates for higher performance (@Shindler09, p 105). But, it may have negative influence if set wrongly. Among the negative influences are potential stress and increased anxiety (of low ranking, fear to fail or technology barrier). In addition, students may invest disproportional time and effort into competition.

Despite some received criticism, a properly set competition can benefit the students greatly. The competition should be relatively short in duration as it consumes a lot of students' energy. Students should be clear about the rules and the goal. They should be properly rewarded and most important, feel that they have reasonable chance to win or achieve high mark [@Shindler09].

@Shelley08 raised the need for more quantitative and statistical analysis of evidence in science education. In this paper we will contribute to this call by offering statistical analysis of the experiment's results.

# Experimental setup

## Data collection

The experiment was conducted during Semester 2 2017. Data was collected during two classes, one at the University of AB (MAST90083), and one at CD University (ETC2420/5242).

## Competition data

Two data sets were compiled for the kaggle challenges: Melbourne property auction prices and spam classification. The Melbourne auction price data was collected by extracting information from real estate auction reports (pdf) collected between Feb 2, 2013 and Dec 17, 2016. The spam classification data was compiled by graduate students at Iowa State University as part of a data mining class, in 2009. Data was compiled by monitoring and extracting information from their emails by class members, over a period of a week, and manually tagging them as spam or ham.

Both data sets were split into training and test sets, for the kaggle challenge. Students had access to the true response variable only for the training data.  For the Melbourne housing data, students were expected to predict price based on the property characteristics. For the spam data, students were expected to build a classifier to predict whether the email as spam or not.

Both data sets are challenging for prediction, with relatively high error rates.

The training and the testing data sets of the Melbourne auction price data were similar but not identical across the two institutions. Some of the variables in the data set were simulated, e.g. property land size and house size. The simulated data was generated slightly differently for different institutions. This was done deliberately to prevent students passing answers from one institution to another.

Students received clean data ready for modelling as we only aimed to examine students ability to apply covered material. If the instructor's aim is rather to mimic real-life situation when prior to modelling/forecasting data needs to be processed and cleaned, then this task should be left to students. In our opinion, this approach is more suitable for advanced classes.

## Participants

MAST90083 is titled Computational Statistics and Data Mining, is designed for postgraduate level, for students with math, statistics, information technology or actuarial backgrounds. It covers modelling both continuous (regression) and categorical (classification) response variables. The 63 students were randomized into one of two kaggle competitions, one focused on regression (R) and the other classification (C). This setup mimics randomized control trials, which is the 'gold standard, in experiment design (@Shelley09CH1, ch. 1).
Students built prediction models and made submissions individually for 16 days, and then were allowed to form groups to compete for another 7 days.

The reason for this strategy was first to motivate each of the students to think about modelling and be actively engaged in the competitions through individual submission. The lecturer allowed to create groups towards the end of the competition to illustrate the advantages of group work and mixing models. Another motivation for this stratagem was the university policy, requiring a strategy to assign students individually in group assignments.

0ne of the 63 students didn't took part in the competition. Another student didn't sit the exam. Therefore, the analysed sample size reduced to 61.

ETC2420/5242, titled Statistical Thinking, covers regression, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the University of AB competition. The 145 undergraduate (2420) students were used as controls for examining performance of the postgraduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete. 
<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

<!--
The data from the follow-up questionnaire was used to examine interest in the competition.

63 enrolled students; postgraduate level; background: math & stats, IT and actuarial science;

Students were randomly assigned to one of two groups, regression (R) or classification (C). Each group will have a data competition designed for that topic, made available through kaggle in class (https://inclass.kaggle.com). Students wored individually to build predictive models, and submit predictions for the first 16 days. On the 16^{th} day of the competition students received an email that they allowed to establish groups and continue the competition as a group. The competition was continued for additional 7 days. In case they chose to form a group, the final score of the group in total was accounted as their individual mark towards the final grade for the subject and analysed in this project.  Only 3 groups were formed, two groups with 3 students and one group with two students. All other students chose to continue the competition individually.

Note that the data competition was compulsory to all the students. The volunteering part was the questionnaire at the end of the semester asking for feedback on aspects of the class such as their competition experience, their favourite topic of the semester, and their enjoyment of the material in the subject.

## Platform

MAST90083 used \url{https://www.kaggle.com/c/amiaspam} and \url{https://www.kaggle.com/c/melbprice}. ETC2420/5242 used \url{https://inclass.kaggle.com/c/vitticeps}. 
-->

## Working with the Kaggle InClass platform

Creating a new competition is surprisingly easy. 
Information on setting up a Kaggle InClass challenge can be found at https://www.kaggle.com/about/inclass/overview. 
To ease the setting process, we created a short video illustrating the steps to establish a competition....

Finding data for a competition may take some time. It is not recommended to take freely available data sets from the Internet. A clever student will always try to check if the data is on the web. If it is, she will use the answers from the data set instead of predicting the answers with a self-developed model. As a statistician, and an instructor in a subject, you or one of your colleagues usually have some interesting data set to use. In such a case, the data set and particularly the target values will be unfamiliar to the students.

Once you open the competition, it runs without any intervention from the instructor. The instructor can monitor students progress: the number of submissions, student scores and even the uploaded data at any time. When the competition ends the "Leaderboard" page provides a list of students ordered by the final score. It also provides all the scores from all past submissions (under "Raw Data" on "Public Leaderboard").

From our experience, we recommend limiting the number of daily submissions to one. It encourages students to think about more efficient improvement of their model before the next submissions.  We also believe that it is better to start the competition in the second half of the semester and finish it two to one week before the end of the semester. It allows to cover many different statistical methods that students may apply during the competition. Finishing some time before the end of the semester reduces students anxiety at the end of the semester and the approaching exams.

# Methodology

## Performance

Better performance is equated to better understanding of the material, as measured in the final exam. MAST90083 and ETC2420/5242 included questions, with several parts, on the final exam related to kaggle challenges. These questions were identified prior to data analysis.

For all questions in the exam, difficulty and discrimination scores were computed, using the mean and standard deviations. Of the questions pre-identified as being relevant to the data challenges, only the parts that corresponded to high level of difficulty and high discrimination were included in the comparison of performance.

Scores for the relevant questions were summed, and converted into percentage of the possible score. The total exam score was converted to a percentage.
One can expect, that on average, student's success rate for each question will be about the same as the success rate in the total exam. Understanding one topic better than another will result in higher success rate for questions asking about the better understood topic compared to the scores for other topics. For example, we would expect from a student with a 70% exam mark to get 70% marks on each of the questions in the exam, if she has similar knowledge level on all the exam topics. If in some topic, say regression, the student has better knowledge, she will perform better on the regression questions. Her success rate on regression question will be higher than 70%. Consequently, her performance on some other questions should be below 70% which is associated with lesser understanding of these topics.

Therefore, performance for each student was computed as the ratio of these two numbers, percentage success in the regression (classification) questions and percentage success in the total exam. A value of 1 would indicate that the student's performance on that set of questions was consistent with their overall exam performance, greater than 1 that they performed better than expected, and lower than 1 meant less than expected on that topic.

Using only the percentage of successes for each set of questions, instead of the proposed ratio, will not differentiate between a better performance and just a better student. Especially in the case of ETC2420/5242 that have a mixed population of master and undergraduate student.

The distribution of the performance scores by group is shown as a boxplot. Focus is on the difference in median between the groups. Permutation tests were conducted to examine difference in median scores for students participating or not in a competition.

## Engagement

The students were allowed to submit at most one prediction per day, while the  competitions were open. The frequency of submissions, and the accuracy (or error) of their predictions, made by individual students, is recorded as a part of the kaggle system. To examine whether engagement improved performance, scores on the questions related to the competition normalised by total exam score (as computed in the performance section) is examined in relation to frequency of submissions during the competition. In addition, performance in the competition as measured by accuracy or error, is also examined in relation to the number of submissions. Scatterplots, correlation and linear models are used to examine the associations.

## Interest

Students in MAST90083 and ETC5242 were invited to give feedback about the course, in particular about the data competitions, before the final exam. This information was voluntary, and students who completed the questionnaire were rewarded with a coupon for a free coffee. The data from this survey was viewed by the researchers after all course grades had been reported.

# Results

## Performance

Figure \ref{fig:MAST90083} shows the data collected in MAST90083. Performance is plotted against type of question, separately for the competition they completed.  The difference in median scores indicates performance improvement. Students who completed the classification competition (left) performed relatively better on the classification questions than the regression questions in the final exam. Conversely, students who participated in the regression competition performed relatively better on the regression questions.

```{r Melb, fig.cap = "\\label{fig:MAST90083} Boxplots of performance on regression and classification questions in the final exam, by type of data competition completed in MAST90083. Students generally performed better on the questions corresponding to the competition they participated in.", fig.height=3, fig.width=6}
library(tidyverse)
library(gridExtra)
library(forcats)
UniMelbData <- read_csv(file="data/deidentified_UMelb.csv") 
UniMelbData <- UniMelbData %>% mutate(Proj_alloc=fct_recode(Proj_alloc, "regression"="Melbourne Price", "classification"="Spam classification"))
UniMelbData <- UniMelbData %>% 
  mutate(perf_class = ((`Q1_p2`+`Q10_p3_A`+`Q10_p7_B`)/12)/(`Total_100`/100)) %>%
  mutate(perf_reg = ((`Q6_p2_A`+`Q6_p2_B`+`Q6_p2_C`+`Q8_p2_C`+`Q8_p2_D`+`Q16_p2_B`)/12)/(`Total_100`/100))

p1 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_class)) +
  geom_boxplot() +
  xlab("Classification competition") +
  ylab("Performance")
p2 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_reg)) +
  geom_boxplot() +
  xlab("Regression competition") +
  ylab("Performance")
grid.arrange(p1, p2, ncol=2)
```

```{r eval=FALSE}
# Didn't take into account difficulty and discrimination
g_pte_class <- ggplot(UniMelbData, aes(Proj_alloc, pte_class)) +
  geom_boxplot() + 
  xlab("") + 
  ylab("CE") + 
  annotate("text", x=0.5, y=0.2, label="A") + 
  ggtitle("Classification Performance")
g_pte_reg <- ggplot(UniMelbData, aes(Proj_alloc, pte_reg)) + 
  geom_boxplot() + 
  xlab("") + 
  ylab("RE") + 
  annotate("text", x=0.5, y=0.4, label="B") + 
  ggtitle("Regression Performance")
g_ptq_class <- ggplot(UniMelbData, aes(Proj_alloc, ptq_class)) +
  geom_boxplot() + 
  xlab("Competition") + 
  ylab("CQ") + 
  annotate("text", x=0.5, y=1.0, label="C")
g_ptq_reg <- ggplot(UniMelbData, aes(Proj_alloc, ptq_reg)) + 
  geom_boxplot() + 
  xlab("Competition") + 
  ylab("RQ") + 
  annotate("text", x=0.5, y=1.0, label="D")
grid.arrange(g_pte_class, g_pte_reg, g_ptq_class, g_ptq_reg, ncol=2)
set.seed(88888)
pdif_C <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_C <- c(pdif_C,
            median(x$pte_class[x$class=="classification"])-
                   median(x$pte_class[x$class=="regression"]))
}
dif_C <- median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="classification"])-
       median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="regression"])
pdif_R <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_R <- c(pdif_R,
            median(x$pte_reg[x$class=="regression"])-                   median(x$pte_reg[x$class=="classification"]))
}
dif_R <- median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="regression"])-
       median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="classification"])
pdif_CQ <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, ptq_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_CQ <- c(pdif_CQ,
            median(x$ptq_class[x$class=="classification"])-
            median(x$ptq_class[x$class=="regression"]))
}
dif_CQ <- median(UniMelbData$ptq_class[UniMelbData$Proj_alloc=="classification"])-
       median(UniMelbData$ptq_class[UniMelbData$Proj_alloc=="regression"])
pdif_RQ <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, ptq_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_RQ <- c(pdif_RQ,
            median(x$ptq_reg[x$class=="regression"])-
              median(x$ptq_reg[x$class=="classification"]))
}
dif_RQ <- median(UniMelbData$ptq_reg[UniMelbData$Proj_alloc=="regression"])-
       median(UniMelbData$ptq_reg[UniMelbData$Proj_alloc=="classification"])
```

```{r eval=FALSE}
# This code was used to examine the scores across all questions, 
# and understand the variability among the groups.
# Difficulty and discrimination for each question is computed. 
UniMelbData_Qs <- UniMelbData %>% 
  select(ID,Q1_p2:Q16_p2_B) %>%
  gather(Q, score, -ID) %>%
  separate(Q, c("Q", "pts", "part")) %>%
  mutate(pts = as.numeric(sub("p", "", pts))) %>%
  mutate(pct = score/pts*100)
diff_discr <- UniMelbData_Qs %>% 
  group_by(Q, part) %>%
  summarise(m = mean(pct, na.rm=TRUE), s = sd(pct, na.rm=TRUE))
diff_discr %>% arrange(m) 
diff_discr %>% arrange(desc(s))
library(plotly)
ggplot(diff_discr, aes(x=m, y=s, label=paste(Q, part))) + geom_point()
ggplotly()
keep <- diff_discr %>% filter(m<71, s>20)
# 1,10, 16B, 6A,B,C, 8C,D, 2,3,4
```

```{r}
# This code conducts the permutation test
pdif_cl <- NULL
pdif_reg <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, perf_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif_cl <- c(pdif_cl,
            median(x$perf_class[x$class=="classification"])-
              median(x$perf_class[x$class=="regression"]))
  pdif_reg <- c(pdif_reg,
            median(x$perf_reg[x$class=="regression"])-
              median(x$perf_reg[x$class=="classification"]))
}
pdif_cl_true <- median(UniMelbData$perf_class[UniMelbData$Proj_alloc=="classification"])-
              median(UniMelbData$perf_class[UniMelbData$Proj_alloc=="regression"])
pdif_reg_true <- median(UniMelbData$perf_reg[UniMelbData$Proj_alloc=="regression"])-
              median(UniMelbData$perf_reg[UniMelbData$Proj_alloc=="classification"])
pval_cl <- length(pdif_cl[pdif_cl > pdif_cl_true])/1000
pval_reg <- length(pdif_reg[pdif_reg > pdif_reg_true])/1000
```

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|r|r|}\hline
Question Set & Median difference & Permutation $p$-value \\\hline
Classification & 0.250 & 0.033\\
Regression & 0.104 & 0.000\\\hline
\end{tabular}
\caption{Comparison of median difference in performance by competition group, for MAST90083 students, using permutation tests. Both sets of medians are significantly different, indicating improved scores for questions on the topic related to the Kaggle competition.}
\end{center}
\label{tab:Melb_Perm}
\end{table}

Table 1 shows the results of permutation testing of median difference between the groups. Generally the results support the competition improved performance. Students who participated in the kaggle challenge for classification scored higher than those that did the regression competition, on the classification problem. Using a permutation test, this corresponds to a significant difference in medians. Similarly the results show that students who did the regression challenge, performed better on these exam questions.

Figure \ref{fig:ETC5242} shows the results for students ETC2420/5242. The boxplots suggest that the students who participated in the challenge performed relatively better than those that didn't on the regression question than expected given their total exam performance.

Only the post-graduate students participated in the regression competition, as their additional assessment requirement. Scores for the question on regression (Q7a,b,c) in the final exam were compared with the total exam score (RE). On these question parts, a, b, c, over all the students all three were in the top 10 of difficulty, with students scoring less than 70%, on average. Parts b, c were in the top 10 for discrimination, and part a was at rank 13.

```{r ETC5242, fig.cap = "\\label{fig:ETC5242} Performance for regression question relative to total exam score for students who did and didn't do the regression data competition in ETC2420/5242.", fig.width=4, fig.height=3}
scores_5242 <- read_csv("data/deidentified_scores_5242.csv")
scores_2420 <- read_csv("data/deidentified_scores_2420.csv")

scores_5242$`Exam Tot` <- apply(scores_5242[,4:27], 1, sum, na.rm=T)
scores_2420$`Exam Tot` <- apply(scores_2420[,4:27], 1, sum, na.rm=T)

scores_5242 <- scores_5242 %>% mutate(perf = ((`7a (10)` + `7b (8)` + `7c (3)`)/21)/(`Exam Tot`/100))
scores_2420 <- scores_2420 %>% mutate(perf = ((`7a (10)` + `7b (8)` + `7c (3)`)/21)/(`Exam Tot`/100))
scores_5242 <- scores_5242 %>% mutate(class="Yes") %>%
  filter(!is.na(perf))
scores_2420 <- scores_2420 %>% mutate(class="No") %>%
  filter(!is.na(perf))

scores <- bind_rows(scores_5242, scores_2420)
ggplot(scores, aes(x=class, y=perf)) + 
  geom_boxplot() + 
  xlab("Participated in competition") +
  ylab("Performance")
```

```{r eval=FALSE}
# This code was used to examine the scores across all questions, and understand 
# the variability in the two groups.
# Difficulty and discrimination for each question is computed. 
# Parts a, b, c of Q7 for ETC2420/5242 are both high in difficulty and discrimination.
# 
scores_5242 <- read_csv("data/deidentified_scores_5242.csv")
scores_2420 <- read_csv("data/deidentified_scores_2420.csv")
poss_marks <- c(4, 2, 2, 2, 3, 3, 9, 8, 5, 2, 2, 3, 3, 3, 2, 3, 3, 6, 10, 8, 3, 3, 3, 2, 4, 2)
scores_pct <- bind_rows(scores_2420, scores_5242)
scores_pct$group <- c(rep("2420", 145), rep("5242", 34))
scores_pct <- scores_pct %>% filter(!is.na(`9b (2)`))
for (i in 1:26)
  scores_pct[,i+1] <- scores_pct[,i+1]/poss_marks[i]

scores_pct_long <- scores_pct %>% gather(q, mark, -group, -id)

# Difficulty
scores_pct_long %>% group_by(q) %>% 
  summarise(m=mean(mark)) %>% 
  arrange(m) %>% print(n=30)
# Discrimination
scores_pct_long %>% group_by(q) %>% 
  summarise(s=sd(mark)) %>% 
  arrange(desc(s)) %>% print(n=30)

scores_pct_m <- scores_pct %>% 
  group_by(group) %>% summarise_all(mean, na.rm=T) %>% select(-id)
scores_pct_s <- scores_pct %>% 
  group_by(group) %>% summarise_all(sd, na.rm=T) %>% select(-id)
scores_pct_s[1,-1] <- scores_pct_s[1,-1]/sqrt(140)*1.96
scores_pct_s[2,-1] <- scores_pct_s[2,-1]/sqrt(34)*1.96

ms <- bind_cols(gather(scores_pct_m, q, value, -group), 
                gather(scores_pct_s, q, value, -group)) %>%
  select(-group1, -q1) %>%
  rename(m=value, s=value1)
ggplot(ms, aes(x=q, y=m, colour=group)) + 
  geom_errorbar(aes(ymin=m-s, ymax=m+s), position="dodge")
```

```{r eval=FALSE}
# This is not the right approach
scores_5242 <- scores_5242 %>% mutate(`adj Exam Tot` = `Exam Tot` - `7b (8)`) %>%
  mutate(`std adj Exam Tot`=(`adj Exam Tot`-mean(`adj Exam Tot`, na.rm=TRUE))/(sd(`adj Exam Tot`, na.rm=TRUE)),
         `std 7b (8)` = (`7b (8)`-mean(`7b (8)`, na.rm=TRUE))/sd(`7b (8)`, na.rm=TRUE))
scores_2420 <- scores_2420 %>% mutate(`adj Exam Tot` = `Exam Tot` - `7b (8)`) %>%
  mutate(`std adj Exam Tot`=(`adj Exam Tot`-mean(`adj Exam Tot`, na.rm=TRUE))/(sd(`adj Exam Tot`, na.rm=TRUE)),
         `std 7b (8)` = (`7b (8)`-mean(`7b (8)`, na.rm=TRUE))/sd(`7b (8)`, na.rm=TRUE))

scores_5242 <- scores_5242 %>% mutate(perf2 = `std 7b (8)` - `std adj Exam Tot`)
scores_2420 <- scores_2420 %>% mutate(perf2 = `std 7b (8)` - `std adj Exam Tot`)
scores_5242 <- scores_5242 %>% mutate(class="Yes") %>%
  filter(!is.na(perf2))
scores_2420 <- scores_2420 %>% mutate(class="No") %>%
  filter(!is.na(perf2))
scores <- bind_rows(scores_5242, scores_2420)

ggplot(scores, aes(x=class, y=perf2)) + 
  geom_boxplot() + 
  xlab("Participated in competiton") +
  ylab("Standardized RE")
```

```{r permtest}
pdif <- NULL
for (i in 1:1000) {
  x <- scores %>% select(class, perf)
  x <- x %>% mutate(class = sample(class))
  pdif <- c(pdif,
            median(x$perf[x$class=="Yes"])-
                   median(x$perf[x$class=="No"]))
}
dif <- median(scores$perf[scores$class=="Yes"])-
       median(scores$perf[scores$class=="No"])
```

Based on the median, the students who participated in the kaggle challenge scored `r round(dif, 2)` higher than those that didn't, a median of `r round(median(scores$perf[scores$class=="Yes"]), 2)` in comparison to `r round(median(scores$perf[scores$class=="No"]), 2)`. Using a permutation test, this corresponds to a significant difference in medians, with $p$-value of `r round(length(pdif[pdif >= dif])/1000, 3)`.

## Engagement

The number of submissions that a student made, may be an indicator of performance on the exam questions related to the competition.  A student who is more engaged in the competition may learn more about the material, and consequently perform better on the exam. Figure \ref{fig:numsubmition} (top row) shows performance on the classification and regression questions, respectively, against their frequency of prediction submissions for the three student groups (MAST90083 classification and regression, ETC5242 regression) competitions. The relationship is weak in all groups, and this mirrors insignificant results from a linear model fit to both subsets. On the other hand, the predictive accuracy improved with the number of submissions for the regression competitions.

The competition performance relative to number of submissions is shown in plots (d)-(f). Each point corresponds to one student, and accuracy or error of the best predictions submitted is used. The regression competition seemed to engage students more than the classification challenge. Students submitted more predictions, and their models improved with more submissions.

<!--
To examine the correlation between student’s engagement levels and the performances in the exam we plot the number of submission during the competition versus the performances in the exam, Figure \ref{fig:numsubmition}. Once again we exam the performances based on two normalizations. Once normalizing by the total possible marks for the relevant cluster of questions ($PTQ$) and once normalizing by total exam marks ($PTE$). For the students participated in the Melbourne Price competition is the cluster of regression questions. For the students participated in the Spam Classification competition is the cluster of questions about classification methods. 
-->

```{r numsubmition, echo = FALSE, fig.cap = "\\label{fig:numsubmition} Scatterplots of the exam performance (a-c) and competition performance (d-f) by number of prediction submissions, for the three student groups. The relationships with exam performance are weak. For the MAST90083 and ETC5242 regression competitions, a clear pattern is that predictions improved substantially with more submissions. (House price in ETC5242 were divided by 100,000, explaining the difference in magnitude of error between two competitions.)" , fig.width=6, fig.height=5}
etc5242 <- read_csv("data/competition_5242.csv")
scores_5242_comp <- scores_5242 %>% left_join(etc5242, by="id")
p1 <- ggplot(filter(UniMelbData, Proj_alloc == "classification"), aes(x=numsub, y=perf_class)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  ylab("Classification performance") + 
  xlab("Num submissions") +
  xlim(c(1,5)) + ggtitle("a. MAST90083 C")
p2 <- ggplot(filter(UniMelbData, Proj_alloc == "regression"), aes(x=numsub, y=perf_reg)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  ylab("Regression performance") + 
  scale_x_continuous("Num submissions", limits=c(1,13), breaks=seq(1, 13, 2)) + 
  ggtitle("b. MAST90083 R")
p3 <- ggplot(scores_5242_comp, aes(x=numsub, y=perf)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  ylab("Regression performance") + 
  xlab("Num submissions") + ggtitle("c. ETC5242")
#grid.arrange(p1, p2, ncol=2)
p4 <- ggplot(filter(UniMelbData, Proj_alloc == "classification"), aes(x=numsub, y=PrivetScore)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  ylab("Accuracy") + 
  xlab("Num submissions") +
  xlim(c(1,5)) + ggtitle("d. MAST90083 C")
p5 <- ggplot(filter(UniMelbData, Proj_alloc == "regression"), aes(x=numsub, y=PrivetScore)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  #ylab("Error") + 
  scale_x_continuous("Num submissions", limits=c(1,13), breaks=seq(1, 13, 2)) + 
  scale_y_continuous("Error ('0000s)", breaks=seq(110000, 170000, 10000), labels=seq(11,17,1)) + ggtitle("e. MAST90083 R")
  #xlim(c(1,12))
p6 <- ggplot(scores_5242_comp, aes(x=numsub, y=score)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) + 
  ylab("Error") + 
  xlab("Num submissions") + ggtitle("f. ETC5242")
grid.arrange(p1, p2, p3, p4, p5, p6, ncol=3)

```

```{r eval=FALSE}
library(broom)
m1 <- lm(perf_class~numsub, data=filter(UniMelbData, Proj_alloc == "classification"))
tidy(m1)
glance(m1)
m2 <- lm(perf_class~numsub, data=filter(UniMelbData, Proj_alloc == "regression"))
tidy(m2)
glance(m2)
```

```{r eval=FALSE}
# Classification: Q1 & Q10 
gsg <- ggplot(UniMelbData %>% filter(Proj_alloc == "Spam classification"), aes(numsub, ptq_class)) + 
  geom_point() + xlab("Num. of submissions") + ylab("")

gse <- ggplot(UniMelbData %>% filter(Proj_alloc == "Spam classification"), aes(numsub, pte_class)) + 
  geom_point() + xlab("Num. of submissions") + ylab("")

grg <- ggplot(UniMelbData %>% filter(Proj_alloc == "Melbourne Price"), aes(numsub, ptq_reg)) + 
  geom_point() + xlab("Num. of submissions") + ylab("")

gre <- ggplot(UniMelbData %>% filter(Proj_alloc == "Melbourne Price"), aes(numsub, pte_reg)) + 
  geom_point() + xlab("Num. of submissions") + ylab("")

pl <- list(grg, gsg,
           gre, gse )

# Create row and column titles
col.titles = c("Question related to project type (PTQ)", "Total exam score (PTE)")
row.titles = c("MElbPrice", "SPAMclass" )

# Add row titles
pl[1:2] = lapply(1:2, function(i) arrangeGrob(pl[[i]], left=row.titles[i]))

# Add column titles and lay out plots
grid.arrange(grobs=lapply(c(1,3), function(i) {
  arrangeGrob(grobs=pl[i:(i+1)], top=col.titles[i/2 + 1], ncol=1)
}), ncol=2)
```

<!--
In Figure \ref{fig:numsubmition} we can clearly see a weak positive correlation between the number of submissions during the data competition and the scores normalized to the total possible marks for the relevant cluster of questions. This suggest that as more engaged the student was with the competition, the question about the methods relevant to her competition were easier to her. There is no correlation between the number of submission and the marks for the relevant cluster of questions normalized by the total exam marks ($PTE$). *BECAUSE … the questions that unrelated to the data competitions (51 points)? students put less effort to learn other material? Harder questions?*

We didn’t found any evidence for correlation between the performances in the competition (final score) and the performances in the exam. This suggest that the single fact of participation improve the students marks in the exam. Not necessarily better students in the competition have grater chances to success in the exam. 
-->

```{r crossperformances, echo=FALSE, fig.cap="\\label{fig:crossperformances} Students performance in classification and regression questions by competition type.", fig.width=6,fig.height=5  }

t <- UniMelbData %>%
 count(Proj_alloc, perf_class >=1, perf_reg >=1, sort = TRUE) %>%
  setNames(., c("Proj_alloc", "outperf_class", "outperf_reg" , "n"))

a <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == FALSE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == FALSE)["n"]) ,")")

b <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]) ,")")

c <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == TRUE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == TRUE)["n"]) ,")")

d <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]) ,")")

a <- "Overall weak students"
b <- "Regression competition \n students perform well\n on regression questions" 
c <- "Overall strong students"
d <- "Classification competition \n students perform well\n on classification questions"

# filter(UniMelbData, Proj_alloc == "regression" , perf_class < 1, perf_reg >=1) %>% select(., perf_class, perf_reg)
# filter(UniMelbData, Proj_alloc == "regression" , perf_class < 1, perf_reg <1) %>% select(., perf_class, perf_reg)

ggplot() +
      annotate("rect", xmin = 0, xmax = 0.99, ymin = 1.01, ymax = 2, fill = "yellow", alpha = 0.25) +
      annotate("rect", xmin = 1.01, xmax = 2, ymin = 0, ymax = 0.99, fill = "yellow", alpha = 0.25) +
      geom_point(data=UniMelbData, aes(x=perf_class, y=perf_reg, color = Proj_alloc), size=3, alpha=0.8) +
      ylab("Regression questions performance") + 
      xlab("Classification questions performance") +
      xlim(0,2) + ylim(0, 2) +
      scale_color_brewer(palette="Dark2", name="Competition", 
                          breaks=c("regression", "classification"),   
                          labels=c("Regression", "Classification")) +
      geom_hline(yintercept = 1, color = "gray", size=0.7) +  #linetype="dotted",
      geom_vline(xintercept = 1, color = "gray", size=0.7) +
      annotate("text", x = c(0.3,0.32,1.7,1.61), y=c(0.1,1.85,1.95,0.18), 
               label = c(  a,  #(1,1)
                           b,  # "(2,2)"
                           c,  #"(3,3)"
                           d )) +
  theme_bw() +
        theme(legend.position="bottom") 
#"(4,4)"
      # annotate("text", x=0.1, y=1.8, label= "(,)")

```

Figure \ref{fig:crossperformances} presents students' scores for classification and regression questions. A score over 1 is considered outperformance (relative to the expectation). Quarters one and three include students that underperform or outperform on both types of questions, respectively. In both cases the number of students that participated in the classification competition is very close to the number of students that participated in the regression competition (excluding a few regression students on the border of score 1). Students in quarters two and four outperform on one type of questions but not on the other type. We can see that more regression students outperform on regression questions than classification students 
(`r filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 12 --> vs. 
 `r filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 7 -->). 
Similarly, classification students do better on classification questions 
(`r filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 11 --> vs. 
 `r filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 3 -->). 
This is another evidence towards positive influence of the data competition on student's performances.

## Interest

<!--
- Show interest statistics. Maybe tables.
- Do topics match the competition they entered for MAST90083
- Match exam performance to engagement?
-->

Figure \ref{fig:likert} shows the survey responses related to the kaggle competition, for MAST90083 and ETC5242. The response rate for MAST90083 was 55%, with 34 of 61 students completing the survey. 
The response rate for ETC5242 was 50%, 17 students out of 34 completed the survey. Overwhelmingly, students reported that they found the competition interesting and helpful for their learning in the course.

After collecting the survey from the students we were told that the questions about students engagement were positively worded. This may have potentially led to some bias. An improved way could be to ask directly about student's engagement, e.g. "How would you rate your level of engagement in this course?" with set answer options of "not at all engaged" -- up to "extremely engaged" with several choices in between. Another improvement could be asking ETC2420 students that didn't take part in the competition about their level of engagement and compare the answers with other students of ETC5242. We acknowledge that the differences in the engagement levels may not necessarily be a result of participation in the competition but it is still an interesting aspect.

The survey was not anonymous. However, the results became available to the lectures only after all the grades were realised to the students. This point was emphasized in the instructions to the students at the beginning of the survey.

```{r likert, fig.cap="\\label{fig:likert}Summary of responses to survey of kaggle competition participants. Overwhelmingly the response to the competition was positive in both classes, especially the questions on enjoyment and engagement in the class, and obtaining practical experience. (Table 2 lists the questions.)"}
MAST90083_survey <- read_xlsx("data/Data Competition Feedback MAST90083.xlsx") %>% filter(`Email Address` != "julia.polak@unimelb.edu.au")
MAST90083_likert <- MAST90083_survey %>% 
  rename(Q1=`1. I found the data competition is great fun.`) %>%
  rename(Q2=`2. Taking part in the data competition contributed a lot to my engagement with the subject.`)  %>%
  rename(Q3=`3. Taking part in the data competition improved my confidence in my understanding of the covered material.`)  %>%
  rename(Q4=`4. Taking part in the data competition improved my confidence in my success in the final exam.`) %>%
  rename(Q5=`5. Taking part in the data competition improved my confidence in my ability to use the acquired knowledge in practical applications.`) %>%
  rename(Q6=`6. I feel that the required time investment in the data competition was worthy.`)
Q1 <- MAST90083_likert %>% count(Q1) %>% filter(!is.na(Q1)) %>%
  mutate(response = fct_relevel(Q1, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q1") %>%
  mutate(pct = n/sum(n))
Q2 <- MAST90083_likert %>% count(Q2) %>% filter(!is.na(Q2)) %>%
  mutate(response = fct_relevel(Q2, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q2")  %>%
  mutate(pct = n/sum(n))
Q3 <- MAST90083_likert %>% count(Q3) %>% filter(!is.na(Q3)) %>%
  mutate(response = fct_relevel(Q3, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q3")  %>%
  mutate(pct = n/sum(n))
Q4 <- MAST90083_likert %>% count(Q4) %>% filter(!is.na(Q4)) %>%
  mutate(response = fct_relevel(Q4, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q4")  %>%
  mutate(pct = n/sum(n))
Q5 <- MAST90083_likert %>% count(Q5) %>% filter(!is.na(Q5)) %>%
  mutate(response = fct_relevel(Q5, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q5")  %>%
  mutate(pct = n/sum(n))
Q6 <- MAST90083_likert %>% count(Q6) %>% filter(!is.na(Q6)) %>%
  mutate(response = fct_relevel(Q6, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q6")  %>%
  mutate(pct = n/sum(n))
MAST90083_Q <- bind_rows(Q1, Q2, Q1, Q2, Q3, Q4, Q5, Q6)

ETC5242_survey <- read_xlsx("data/Data Competition Feedback ETC Di.xlsx") %>% filter(Name != "Dianne Cook")
ETC5242_likert <- ETC5242_survey %>% 
  rename(Q1=`1. I found the data competition is great fun.`) %>%
  rename(Q2=`2. Taking part in the data competition contributed a lot to my engagement with the subject.`)  %>%
  rename(Q3=`3. Taking part in the data competition improved my confidence in my understanding of the covered material.`)  %>%
  rename(Q4=`4. Taking part in the data competition improved my confidence in my success in the final exam.`) %>%
  rename(Q5=`5. Taking part in the data competition improved my confidence in my ability to use the acquired knowledge in practical applications.`) %>%
  rename(Q6=`6. I feel that the required time investment in the data competition was worthy.`)
Q1 <- ETC5242_likert %>% count(Q1) %>% filter(!is.na(Q1)) %>%
  mutate(response = fct_relevel(Q1, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q1") %>%
  mutate(pct = n/sum(n))
Q2 <- ETC5242_likert %>% count(Q2) %>% filter(!is.na(Q2)) %>%
  mutate(response = fct_relevel(Q2, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q2")  %>%
  mutate(pct = n/sum(n))
Q3 <- ETC5242_likert %>% count(Q3) %>% filter(!is.na(Q3)) %>%
  mutate(response = fct_relevel(Q3, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q3")  %>%
  mutate(pct = n/sum(n))
Q4 <- ETC5242_likert %>% count(Q4) %>% filter(!is.na(Q4)) %>%
  mutate(response = fct_relevel(Q4, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q4")  %>%
  mutate(pct = n/sum(n))
Q5 <- ETC5242_likert %>% count(Q5) %>% filter(!is.na(Q5)) %>%
  mutate(response = fct_relevel(Q5, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q5")  %>%
  mutate(pct = n/sum(n))
Q6 <- ETC5242_likert %>% count(Q6) %>% filter(!is.na(Q6)) %>%
  mutate(response = fct_relevel(Q6, "Strongly Disagree", "Disagree", "No Strong Feelings", "Agree", "Strongly Agree")) %>% 
  mutate(response = fct_recode(response, "Neutral" = "No Strong Feelings")) %>%
  mutate(question = "Q6")  %>%
  mutate(pct = n/sum(n))
ETC5242_Q <- bind_rows(Q1, Q2, Q1, Q2, Q3, Q4, Q5, Q6)
MAST90083_Q <- MAST90083_Q %>% mutate(class = "MAST90083")
ETC5242_Q <- ETC5242_Q %>% mutate(class = "ETC5242")
Q <- bind_rows(MAST90083_Q, ETC5242_Q)
Q <- Q %>% mutate(class = fct_relevel(class, "MAST90083", "ETC5242"))
ggplot(Q, aes(x=question, weight = n, 
              fill=fct_relevel(response, "Strongly Disagree", 
                               "Disagree", "Neutral", "Agree", 
                               "Strongly Agree"))) + 
  xlab("") + ylab("") +
  facet_wrap(~class, ncol=2) +
  geom_bar(position="fill") + coord_flip() +
  scale_fill_manual("", values=c("#CA0020","#F4A582","#F7F7F7","#92C5DE", "#0571B0")) + 
  theme_bw() +
  theme(legend.position = "bottom") 
```

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|p{15cm}|}\hline
Label & Question \\\hline
Q1 & I found the data competition is great fun.\\\hline
Q2 & Taking part in the data competition contributed a lot to my engagement with the subject.\\\hline
Q3 & Taking part in the data competition improved my confidence in my understanding of the covered material.\\\hline
Q4 & Taking part in the data competition improved my confidence in my success in the final exam.\\\hline
Q5 & Taking part in the data competition improved my confidence in my ability to use the acquired knowledge in practical applications.\\\hline
Q6 & I feel that the required time investment in the data competition was worthy. \\\hline
\end{tabular}
\end{center}
\caption{Questions asked in the survey of competition participants.}
\label{survey_Q}
\end{table}

# Discussion

This paper has described the setup and results of an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The experiment was conducted in the classroom settings as part of the normal teaching of the courses, which imposes limitations on the design. However, with both MAST90083 and ETC2420/5242, either a randomized assignment of students to two topic groups, or a control group was possible, which enabled comparing performance.

The primary finding is that participating in competitions results produces a statistically significant improvement in the learning of the topic, although the effect size is small.  Secondarily, anecdotally, the competitions enhanced interest and engagement in the course.

# Acknowledgments

This project (title: Effect of Data Competition on Learning Experience) has been approved by the Faculty of Science Human Ethics Advisory Group University of AB (ID: 1749858.1 on September 4, 2017) and by CD University Human Research Ethics Committee (ID: 9985 on August 24, 2017).

This document was produced in R [@R] with the package knitr [@knitr]. Data cleaning was conducted using tidyr [@tidyr], dplyr [@dplyr] and plots were made with ggplot2 [@ggplot2]. The materials to reproduce the work are available at https://github.com/XXX.

<!--
https://github.com/dicook/paper-quoll. 
-->
Information on setting up a Kaggle InClass challenge can be found at https://www.kaggle.com/about/inclass/overview.